---
output:
  github_document:
    html_preview: TRUE
---

```{r readme-01, echo = FALSE}
knitr::opts_chunk$set(
  fig.asp = 2/3,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# yoctopuce

<!-- badges: start -->
[![gginnards status badge](https://aphalo.r-universe.dev/badges/yoctopuce)](https://aphalo.r-universe.dev/yoctopuce)
[![web site](https://img.shields.io/badge/documentation-yoctopuce-informational.svg)](https://docs.r4photobiology.info/yoctopuce/)
<!-- badges: end -->

## Purpose

[Yoctopuce](https://www.yoctopuce.com/) makes electrically isolated USB modules and special hubs suitable for diverse automation tasks, including data acquisition and logging, and various actuators suitable for control. Yoctopuce supplies libraries for several computing languages but not for R. Using the _yoctopuce_ Python library from R is rather easy thanks to R package ['reticulate'](https://rstudio.github.io/reticulate/). R package 'yoctopuce' makes it easier for R users unfamiliar with package 'reticulate' and/or Python to use this library. Package 'yoctopuce' also includes an article (published only as part of the online documentation) with use case examples for different Yoctopuce USB modules.

## Use

If using a virtual hub instead of a hardware one, the [VirtualHub software](https://www.yoctopuce.com/EN/virtualhub.php) should be installed and running on the computer or microcontroller board where the USB modules are plugged in.

```{r, eval=FALSE}
library(yoctopuce)
```
Python and the Python 'yoctopuce' library have to be installed on the computer where the R code will run. The library is installed in a separate Python environment `r-yoctopuce', that is used only by R package 'yoctopuce'. This is to avoid interfering with other local uses of Python. Installation, of course, needs to be done only once.

```{r, eval=FALSE}
install_python()
install_yoctopuce()
```

On each R session where communication with Yoctopuce modules is needed, the library needs to be imported and registered. Based on the USB module being targeted we import Python library modules to make them available in R and register the hub (virtual or hardware) that will be used to connect to the modules. The example below assumes we will use a **YoctoRelay** and a **YoctoMeteo** module through a local instance of the virtual hub.

See the ['yoctopuce' Python library documentation](https://www.yoctopuce.com/EN/doc/reference/yoctolib-php-EN.html) for details. Be aware that the imports into R are done by `"function"` name, not by USB module name. Some USB modules support more than one "function" and multiple USB modules that are logically equivalent normally support the same "function". Say for both YoctoRelay and YoctoPowerRelay USB modules the import is the same `yocto_relay`, while for YoctoMeteo with its different "functions", three imports are needed to be able to use all of them, as shown in the example below. Additional imports would be needed to access the data logger built into the YoctoMeteo module.

```{r, eval=FALSE}
y_initialise("yocto_relay", "yocto_humidity", "yocto_temperature", "yocto_pressure")
ls(pattern = "^yocto")
```

The first line of code above creates R objects `yocto_api`, `yocto_humidity`, `yocto_temperature`, `yocto_pressure` giving access to the functions and objects from the Python library using `$` notation. However, one needs first to find the module one intends to use and create a wrapper object. The serial number plus "function" name within the module or an user-assigned "logical" name of the module "function", is used to locate it. Here we use `"RELAY1"`, the name of the first of two relays ("functions") in our YoctoRelay module. When using default names, only one module of a given type can be used. With multiple identical modules, either names should be changed to unique ones, or serial numbers used. Using default names makes the code usable unchanged with any YoctoPuce module of a given type, while using different names or serial numbers makes it possible to individually access multiple modules of the same type through the same hub.

```{r, eval=FALSE}
Relay1 <- yocto_relay$YRelay$FindRelay("RELAY1")
# in case of failure the Python error message is displayed

# we can check the serial number of the found relay
Relay1$describe()
Relay1$get_functionId()

# We can then, for example, flip the relay switch to ON ("B") state for 200 ms
Relay1$pulse(200)
```

## YoctoPuce documentation

The [YoctoPuce website](https://www.yoctopuce.com/) provides many code examples in module manuals, tutorials and in blog posts. The libraries for different computer languages use consistent naming and design and code examples using versions of the 'yoctopuce' other than the Python one are easy to translate into the Python equivalent. Except for functions `y_initialise()` and `install_yoctopuce()` you will be using the functions from the Python library calling them through the "bridge" provided. As no wrapper R functions are used, at least in principle, R package 'yoctopuce' should give always access to all the functions available in the installed version of the yocotpuce Python library, even in the future.

## Package 'reticulate'

Currently, package 'reticulate' is imported in whole and re-exported. Thus, when package 'yoctopuce' is loaded and attached, package 'reticulate' is also loaded and attached. So, functions and objects from 'reticulate' can be accessed directly by their name, as shown above for `install_phyton()`.

## Installation

To use this package you will need to have at least one USB module from Yoctopuce. To access the module you will need an active Yocto hub. This can be a virtual (software hosted in the local or another computer) or a hardware Yocto Hub from Yoctopuce. Virtual hubs must be running, either started manually or automatically at computer start-up. The virtual hub can be installed as a background process or run as a regular application. In some infrequent cases access to the local USB ports from within the same R using other drivers can interfere with the access using R package 'yoctopuce'. In such cases running the virtual hub as a background process can sometimes help.

Installation of the current unstable version from a CRAN-like repository at R-Universe (Binaries for Mac, Win, Webassembly, and Linux, as well as sources available) is possible by explicitly listing the repository (as shown) or by adding the `https://aphalo.r-universe.dev` repository to the repositories known to R:

```{r universe-installation, eval=FALSE}
install.packages('yoctopuce', 
                 repos = c('https://aphalo.r-universe.dev', 
                           'https://cloud.r-project.org'))
```

Installation of the current unstable version from GitHub (from sources) is also possible from the `main` branch (as shown), but unnecessary unless a branch different from `main` or an older commit is the target:

```{r gh-instalaltion, eval=FALSE}
# install.packages("devtools")
devtools::install_github("aphalo/yoctopuce")
```

## Documentation

HTML documentation is available at (https://docs.r4photobiology.info/yoctopuce/), including one vignette and one article.

News about updates are regularly posted at (https://www.r4photobiology.info/).

## Contributing

Please report bugs and request new features at (https://github.com/aphalo/yoctopuce/issues). Pull requests are welcome at (https://github.com/aphalo/yoctopuce).

## Citation

If you use this package to produce scientific or commercial publications, please cite according to:

```{r}
citation("yoctopuce")
```

## License

© 2024 Pedro J. Aphalo (pedro.aphalo@helsinki.fi). Released under the GPL, version 2 or greater. This software carries no warranty of any kind.

