---
title: "TSL2591 I2C lux sensor"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

This is a demonstration of how to use an I2C light sensor connected to a
Yosto-I2C module directly from R. We use the _job_ file provided by Yoctopuce
in the article
[Yocto-I2C and TSL2591](https://www.yoctopuce.com/EN/article/yocto-i2c-and-tsl2591).

We will use two R packages unrelated to the interaction with the USB module.

```{r}
library(readr)
library(ggplot2)
```

We use R package 'yoctopuce' to get access to the Python version of of the
[Yoctopuce library](https://www.yoctopuce.com/EN/libraries.php) initializing 
the modules from the library that we will use.

```{r}
library(yoctopuce)

init_yoctopuce("yocto_api", 
               "yocto_i2cport", 
               "yocto_files", 
               "yocto_genericsensor")
```
We upload the text file `TSL2591.job` with the code for the job to be run in the 
Yocto-I2C module. The file can be downloaded from [Yoctopuce's files archive](https://www.yoctopuce.com/pubarchive/2020-11/Yocto-I2C-TSL2591_1.zip).
However, here I use a local copy previously downloaded to my computer.

The call to `gsub()` fixes the file end-of-line encoding in case it follows
Windows conventions instead of UNIX/Linux ones. It also seems necessary in other
cases as it forces the conversion from a Python object into an R `character`
vector.

```{r}
job_text <- read_file("./Yocto-I2C-TSL2591_1/TSL2591.job") |>
  gsub("\\r\\n", "\n", x = _)

file_system <- yocto_files$YFiles$FindFiles("YI2CMK01-1214E2.files")

file_system$isOnline()
file_system$upload("TSL2591.job", job_text)
```

To see the job file's text we can use `cat()` as it decodes the embedded
character scapes.

```{r}
cat(job_text)
```

The Yocot-I2C module plays the role of I2C master
and the TSL2591 breakout board the role of an I2C slave. 
We set I2C-communication related parameters in the  module.

```{r}
I2C <- yocto_i2cport$YI2cPort$FindI2cPort("YI2CMK01-1214E2")
if (is.null(I2C)) {
  stop('No module connected (check cable)')
}

I2C$isOnline()
I2C$describe()

# settings
I2C$set_i2cMode("100kbps,2000ms,NoRestart")
I2C$set_i2cVoltageLevel(I2C$I2CVOLTAGELEVEL_3V3)
I2C$set_protocol("Line")
# I2C$saveToFlash() # making settings permanent, run only if needed!

# checks
I2C$get_i2cMode()
I2C$get_i2cVoltageLevel()
I2C$get_protocol()
```
Activate the job and set it to run on module at start up.

```{r}
I2C$get_currentJob()
I2C$selectJob("TSL2591.job")
I2C$set_startupJob("TSL2591.job")
I2C$get_currentJob()
I2C$get_startupJob()
```

We use only one of the 10 generic sensors in the Yocto_I2C module, 
`genericSensor1` as set in the script.

```{r}
illum.sensor <- yocto_genericsensor$YGenericSensor$FindGenericSensor("YI2CMK01-1214E2.genericSensor1")
if (is.null(illum.sensor)) {
  stop('No module connected (check cable)')
}

illum.sensor$describe()
illum.sensor$get_resolution()
illum.sensor$get_valueRange()

illum.sensor$get_advertisedValue()
illum.sensor$get_highestValue()
illum.sensor$get_lowestValue()
illum.sensor$get_currentRawValue()
```

Each generic or specific sensor has a data logger function. Here we access
the datalogger. We start by making sure that it is enabled and the memory
free of old data.

```{r}
illum.logger <- illum.sensor$get_dataLogger()
illum.logger$describe()
#illum.logger$set_clearHistory()
illum.logger$set_autoStart(illum.logger$AUTOSTART_ON)
illum.logger$set_recording(illum.logger$RECORDING_ON)
# I2C$saveToFlash() # making settings permanent, run only if needed!

as.POSIXct(illum.logger$get_timeUTC(), origin = "1970-01-01")
illum.logger$get_usage() # as % of logger memory
```
```{r}
# Enumerate all connected sensors
sensor <- illum.sensor$FirstSensor()

illum.dataset <- yocto_api$YConsolidatedDataSet(0, 0, sensor)
# record <- list()
# illum.dataset$nextRecord(record)
```

```{r}
py_run_string("
# Enumerate all connected sensors
sensorList = []
sensor = YSensor.FirstSensor()
while sensor is not None:
    sensorList.append(sensor)
    sensor = sensor.nextSensor()

# Generate consolidated CSV output for all sensors
data = YConsolidatedDataSet(0, 0, sensorList)
record = []
while data.nextRecord(record):
    line = datetime.datetime.fromtimestamp(record[0]).isoformat()
    for idx in range(1, len(record)):
        line += \";%.3f\" % record[idx]
    print(line)")
```


```{r}
light.df <- read_delim("data.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)
light.df <- light.df[ , 2:3]
```

```{r}
colnames(light.df) <- c("time", "illuminance")
ggplot(light.df, aes(time, illuminance)) +
  geom_line()
```


