---
title: "TSL2591 I2C lux sensor"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: html_document
---

This is a demonstration of how to use the TSL2591 I2C light sensor connected to
a Yosto-I2C module directly from R. We use an edited version of the _job_ file
provided by Yoctopuce in the article [Yocto-I2C and
TSL2591](https://www.yoctopuce.com/EN/article/yocto-i2c-and-tsl2591).

The original job code used an integration time of 100 ms, while in the editted
job we use 500 ms. The intrgration time and gain are set together with the
same command. The last digit sets the integration time, while the second last
sets the gain. I changed `{"writeLine": "{S}52A100{P}"}` into 
`{"writeLine": "{S}52A104{P}"}`. and simlarly for the other gain settings.

The original job code read only channel 0 (CH0) of the TSL2591, the editted
script reads also channel 1 (CH1) of TSL2591. CHO is sensitive to VIS + IR
radiation, while CH1 is sensitive only to IR. To obtain a reading of
illuminance valid under different light sources, the readings from CH1 are
used to correct for the IR sensirtivity of CH0.

The original job code stored the scaled counts from CH1 in GenericSensor1
directly. The editted code saves scaled counts for CHO and CH1 in variables,
and then sets GenericSensor1 and GenericSensor2 with them. The corrected
"light" value is calculated from the two variables and assigned to 
GenericSensor3.

We will use two R packages unrelated to the interaction with the USB module.

```{r}
library(readr)
library(ggplot2)
```

We use R package 'yoctopuce' to get access to the Python version of of the
[Yoctopuce library](https://www.yoctopuce.com/EN/libraries.php) initializing 
the modules from the library that we will use.

```{r}
library(yoctopuce)

init_yoctopuce("yocto_api", 
               "yocto_i2cport", 
               "yocto_files", 
               "yocto_genericsensor")
```
We upload the text file `TSL2591-2ch.job` with the code for the job to be run in the 
Yocto-I2C module. The file will be included in the R package 'yocotopuce'.
However, here I use a local copy previously downloaded to my computer.

The call to `gsub()` fixes the file end-of-line encoding in case it follows
Windows conventions instead of UNIX/Linux ones. It also seems necessary in other
cases as it forces the conversion from a Python object into an R `character`
vector.

```{r}
job_text <- read_file("./Yocto-I2C-TSL2591_1/TSL2591-2ch.job") |>
  gsub("\\r\\n", "\n", x = _)

file_system <- yocto_files$YFiles$FindFiles("YI2CMK01-1214E2.files")

file_system$isOnline()
file_system$upload("TSL2591.job", job_text)
```

To see the job file's text we can use `cat()` as it decodes the embedded
character scapes.

```{r}
cat(job_text)
```

The Yocot-I2C module plays the role of I2C master
and the TSL2591 breakout board the role of an I2C slave. 
We set I2C-communication related parameters in the  module.

```{r}
I2C <- yocto_i2cport$YI2cPort$FindI2cPort("YI2CMK01-1214E2")
if (is.null(I2C)) {
  stop('No module connected (check cable)')
}

I2C$isOnline()
I2C$describe()

# settings
I2C$set_i2cMode("100kbps,2000ms,NoRestart")
I2C$set_i2cVoltageLevel(I2C$I2CVOLTAGELEVEL_3V3)
I2C$set_protocol("Line")
# I2C$saveToFlash() # making settings permanent, run only if needed!

# checks
I2C$get_i2cMode()
I2C$get_i2cVoltageLevel()
I2C$get_protocol()
```
Activate the job and set it to run on module at start up.

```{r}
I2C$get_currentJob()
I2C$selectJob("TSL2591.job")
I2C$set_startupJob("TSL2591.job")
I2C$get_currentJob()
I2C$get_startupJob()
```

We use three of the 10 generic sensors in the Yocto_I2C module, 
`genericSensor1` as set in the script.

```{r}
VIS.sensor <- yocto_genericsensor$YGenericSensor$FindGenericSensor("YI2CMK01-1214E2.genericSensor1")
if (is.null(VIS.sensor)) {
  stop('No module connected (check cable)')
}
IR.sensor <- yocto_genericsensor$YGenericSensor$FindGenericSensor("YI2CMK01-1214E2.genericSensor2")
Illum.sensor <- yocto_genericsensor$YGenericSensor$FindGenericSensor("YI2CMK01-1214E2.genericSensor3")

VIS.sensor$describe()
IR.sensor$describe()
Illum.sensor$describe()

Illum.sensor$get_resolution()
Illum.sensor$get_valueRange()

Illum.sensor$get_advertisedValue()
Illum.sensor$get_highestValue()
Illum.sensor$get_lowestValue()
Illum.sensor$get_currentRawValue()
```

Each generic or specific sensor has a data logger function. Here we access
the datalogger. We start by making sure that it is enabled and the memory
free of old data.

```{r}
Illum.logger <- Illum.sensor$get_dataLogger()
Illum.logger$describe()
#illum.logger$set_clearHistory()
Illum.logger$set_autoStart(illum.logger$AUTOSTART_ON)
Illum.logger$set_recording(illum.logger$RECORDING_ON)
# I2C$saveToFlash() # making settings permanent, run only if needed!

as.POSIXct(Illum.logger$get_timeUTC(), origin = "1970-01-01")
Illum.logger$get_usage() # as % of logger memory
```
```{r}
# Enumerate all connected sensors
sensor <- Illum.sensor$FirstSensor()

Illum.dataset <- yocto_api$YConsolidatedDataSet(0, 0, sensor)
# record <- list()
# Illum.dataset$nextRecord(record)
```

```{r, eval=FALSE}
# does not work
py_run_string("
# Enumerate all connected sensors
sensorList = []
sensor = YSensor.FirstSensor()
while sensor is not None:
    sensorList.append(sensor)
    sensor = sensor.nextSensor()

# Generate consolidated CSV output for all sensors
data = YConsolidatedDataSet(0, 0, sensorList)
record = []
while data.nextRecord(record):
    line = datetime.datetime.fromtimestamp(record[0]).isoformat()
    for idx in range(1, len(record)):
        line += \";%.3f\" % record[idx]
    print(line)")
```


```{r}
data3 <- read_delim("inst-not/TSL2591-Lux-I2C-sensor/data-2ch-3log.csv", 
                    delim = ";", escape_double = FALSE, trim_ws = TRUE,
                    col_select = 2:5)
colnames(data3) <- c("time", "VIS", "IR", "Illum")
data3 <- subset(data3, time > ymd_hm("2024-09-29 15:00"))
```

```{r}
ggplot(data3) + 
  geom_line(aes(time, VIS)) +
  geom_line(aes(time, IR), colour = "red") +
  expand_limits(y = 0)
  
ggplot(data3, aes(time, Illum)) + 
  geom_line() +
  expand_limits(y = 0)

ggplot(data3, aes(time, VIS-IR)) + 
  geom_line() +
  expand_limits(y = 0)
```


